Towards a Pandemic-Scale Ancestral Recombination Graph of SARS-CoV-2

### 2023-04-13 paper.tex
Some analysis of the pango recombinants

\begin{figure} \centering
\includegraphics[width=\textwidth]{figures/cophylogeny_wide.pdf}
\caption{\label{fig:cophylogeny} Tanglegram comparing the first local tree in the
wide ARG (sampled to mid-2021) and an “all-time” global NextStrain tree (sampled
to January 21, 2023). The trees were pruned down to the samples present in both
the trees. Light grey lines match the corresponding samples between the two
trees; black circles indicate identical sample partitions between the two trees.
Terminal branches are colour-coded according to the Pango lineage status
assigned to the tip samples; no Pango recombinants are shown, as none are shared
between the trees. The tanglegram was generated with the aid of the Neighbor-Net
algorithm \citep{Scornavacca2011-mg} implemented in Dendroscope
version 3.8.5 \citep{Huson2012-ys}. See Figure~\ref{fig:cophlogeny_long} for the equivalent cophylogeny
for the Long ARG.}
\end{figure}

@@

\begin{figure}
\begin{tabularx}{\textwidth}{cc}
\includegraphics[width=0.4\textwidth]{figures/Pango_XA_nxcld_tight_graph.pdf}
&
\begin{tabular}[b]{llr}
Lineage & number of samples & other details... \\
XA & & \\
XF & & \\
XG & & \\
XH & & \\
XL & & \\
XR & & \\
XT & & \\
XW & & \\
XY & & \\
XAC & & \\
XAE & & \\
XAK & & \\
XAA & & \\
XAB & & \\
XAG & & \\
\end{tabular}\\
\multicolumn{2}{c}{
\includegraphics[width=\textwidth]{figures/Pango_XAG_nxcld_tight_graph.pdf}}
\end{tabularx}
\caption{Single origin Pango X lineages. (A) Subgraph for XA (Wide ARG).
(B) Details for the the single-origin Pango
X lineages in the long ARG. (C) XAA, XAB and XAG share a single recombinant
origin in the Long ARG. }
\end{figure}

The sequences of 14 Pango recombinants trace back to a single recombination
node. The details are given in Figure X, where we [give the details of
the numbers of samples, etc for each of these 14 lineages] and
show the lineage subgraph for X?, which makes this simple structure
clear. [Explan the lineage subgraph format a bit]

@@

\begin{figure} \centering
\includegraphics[width=\textwidth]{figures/supp_cophylogeny_long.pdf}
\caption{\label{fig:cophlogeny_long}Tanglegram equivalent to that in
Figure~{fig:cophlogeny}, but for the long ARG (i.e. subsampled to mid-2022). }
\end{figure}

>>>>

\begin{figure} \centering
\includegraphics[width=\textwidth]{figures/cophylogeny_wide.pdf}
\caption{\label{fig:cophylogeny} Tanglegram comparing the first local tree in the
wide ARG (sampled to mid-2021) and an “all-time” global NextStrain tree (sampled
to January 21, 2023). The trees were pruned down to the samples present in both
the trees. Light grey lines match the corresponding samples between the two
trees; black circles indicate identical sample partitions between the two trees.
Terminal branches are colour-coded according to the Pango lineage status
assigned to the tip samples; no Pango recombinants are shown, as none are shared
between the trees. The tanglegram was generated with the aid of the Neighbor-Net
algorithm \citep{Scornavacca2011-mg} implemented in Dendroscope
version 3.8.5 \citep{Huson2012-ys}. See
Figure~\ref{fig:cophylogeny_long} for the equivalent cophylogeny
for the Long ARG.}
\end{figure}

@@

\begin{table}
% \begin{footnotesize}
\begin{tabular}{lllrl}
\toprule
{} &         parent\_lineages &          pango\_parents &   n &
descendant\_lineages \\
lineage &                         &                        &     &
\\
\midrule
XA      &   [B.1.177.18, B.1.1.7] &     [B.1.1.7, B.1.177] &   5 & \{'XA': 5\} \\
XAA     &          [BA.1, BA.2.9] &         [BA.1*, BA.2*] &   8 &  \{'XAB':
39, 'XAG': 17, 'XAA': 8, 'XQ': 2, 'XU'... \\
XAC     &     [BA.2.3, BA.1.17.2] &  [BA.2*, BA.1*, BA.2*] &  27 &
\{'XAC': 27\} \\
XAE     &            [BA.2, BA.1] &         [BA.2*, BA.1*] &   9 &
\{'XAE': 9\} \\
XAG     &          [BA.1, BA.2.9] &         [BA.1*, BA.2*] &  17 &  \{'XAB':
39, 'XAG': 17, 'XAA': 8, 'XQ': 2, 'XU'... \\
XF      &            [AY.4, BA.1] &    [B.1.617.2*, BA.1*] &   2 &
\{'XF': 2\} \\
XG      &         [BA.1.17, BA.2] &         [BA.1*, BA.2*] &  32 &
\{'XG': 32, 'XAB': 1\} \\
XH      &       [BA.1.20, BA.2.9] &         [BA.1*, BA.2*] &  11 &  \{'XAF':
34, 'XH': 11, 'B.1.1.529': 3, 'XE': 3,... \\
XL      &       [BA.1.17.2, BA.2] &         [BA.1*, BA.2*] &  10 &
\{'XL': 10, 'XAB': 1, 'XU': 1, 'XQ': 1\} \\
XR      &          [BA.1.1, BA.2] &       [BA.1.1*, BA.2*] &   8 &
\{'XQ': 9, 'XR': 8, 'XAB': 2\} \\
XT      &  [BA.2.23, Unknown (R)] &         [BA.2*, BA.1*] &   1 &  \{'BA.2':
1, 'BA.1': 1, 'BA.1.9': 1, 'BA.2.23':... \\
XW      &       [BA.1.1.15, BA.2] &         [BA.1*, BA.2*] &  11 &
\{'XW': 11, 'XN': 2\} \\
XY      &          [BA.1.1, BA.2] &         [BA.1*, BA.2*] &   6 &
\{'XY': 6, 'XAF': 1\} \\
\bottomrule
\end{tabular}
\caption{\label{tab:pango-single-origin}
Table summarising the status of the Pango X-lineages in the
Long ARG that are inferred to have a single recombinant origin
based on the NextClade Pango lineage assignments. Imputed parent lineages
are shown from the ARG and also to parent lineages as defined by
the official Pango designation alias key.
FIXME: if we like this table then we can reformat it so that lists
at the end are spread over multiple lines rather than using Python
dict notation }
\end{table}

\begin{figure}
\begin{tabularx}{\textwidth}{c}

% \end{footnotesize}
% \\

\includegraphics[width=0.4\textwidth]{figures/Pango_XA_nxcld_tight_graph.pdf}
\\ % &
% \multicolumn{2}{c}{
\includegraphics[width=\textwidth]{figures/Pango_XAG_nxcld_tight_graph.pdf}
% }
\end{tabularx}
\caption{\label{fig:pango-single-origin-graph}
Single origin Pango X lineages. (A) Subgraph for XA (Wide ARG).
(B) Details for the the single-origin Pango
X lineages in the long ARG. (C) XAA, XAB and XAG share a single recombinant
origin in the Long ARG. }
\end{figure}

The sequences of 13 Pango recombinant lineages trace back to a single recombination
node, as summarised in Table~\ref{tab:pango-single-origin}. We can
see that XA, XAC, XAE, and XF form monophyletic clades, in which all of the
samples given that Pango designation descendand from a single recombination
node.

The details are given in Figure X, where we [give the details of
the numbers of samples, etc for each of these 14 lineages] and
show the lineage subgraph for X?, which makes this simple structure
clear. [Explan the lineage subgraph format a bit]

We exclude XAK it only has a single sample in this dataset, and
that sample traces to an older recombination node that
that is likely to be spurious since it has a very large number of descendants.

@@

\begin{figure} \centering
\includegraphics[width=\textwidth]{figures/supp_cophylogeny_long.pdf}
\caption{\label{fig:cophylogeny_long}Tanglegram equivalent to that in
Figure~\ref{fig:cophylogeny}, but for the long ARG (i.e. subsampled to mid-2022). }
\end{figure}
